# Future Model Comparisons

```{r}
#| label: setup
#| include: false

library(tidyverse)
library(kableExtra)
library(scales)

format_los_comp_kbl <- function(los_comp) {
  los_comp %>%
  mutate(Intersection = ifelse(
    Intersection == "SB Interchange",
    "SB Interchange*",
    Intersection
  )) %>% 
  kbl(align = "c", booktabs = TRUE, linesep = "") %>% 
  kable_styling(latex_options = "scale_down") %>% 
  column_spec(column = 4:ncol(final_am_los_comp_formatted), width = "3cm") %>% 
  collapse_rows(columns = c(1,2), valign = "middle", target = 1) %>% 
  footnote(symbol = "This intersection was signalized in the existing model, but is unsignalized in the build model.")
}
```

## Introduction

MATSim modeling software roughly predicted future growth in the Payson area.
The team used a growth rate of 30% over the next 27 years to predict impacts on the existing and built models.
This section outlines the process used to run and compare the models and reviews the results of the models.

## Methodology

To simulate future growth in the region, the team multiplied existing 2023 traffic volumes by a factor of 1.30.
VISSIM models then simulated future traffic conditions.
The team implemented the additional modifications listed below to simulate future situations:

-   Signal timings were adjusted using Synchro (source?) to provide optimal cycle lengths and split timings.
    The team used a 150-second cycle length in all signals and optimized splits under existing signal characteristics.

-   The team modified the build model to allow double lane circulation on the SB interchange between the off and on-ramps.
    The prior model showed low delays on the SB approach, to the detriment of the off-ramp, as the off-ramp has much greater demand than the SB approach.

-   Priority rules were added to the PM no-build model to prevent vehicles from blocking the intersections to allow for cross-movement travel.
    The team collected travel time and nodal results like in previous chapters.
    This analysis will not include a model with current volumes and the build model.

## Results

Tables [-@tbl-am-los-comp_build] and [-@tbl-pm-los-comp_build] show the results Build model in the AM and PM peak periods, compared with the No-Build scenario model.

```{r}
#| label: tbl-am-los-comp_build
#| tbl-cap: Comparison of LOS Results (AM)

targets::tar_load(final_am_los_comp_formatted)

final_am_los_comp_formatted %>% format_los_comp_kbl()
```

```{r}
#| label: tbl-pm-los-comp_build
#| tbl-cap: Comparison of LOS Results (PM)

targets::tar_load(final_pm_los_comp_formatted)

final_pm_los_comp_formatted %>% format_los_comp_kbl()
```

The travel times for each vehicle and each model run were then averaged and summarized by movement.
@fig-tt_build compares average travel times from the Build model with the No-Build model for both AM and PM peak periods.
However, travel distances are also increased due to the roundabout design, and the increase in travel times is roughly proportional to this increase in distance.


```{r}
#| label: fig-tt_build
#| fig-cap: Travel time comparison between existing and build models.

targets::tar_load(nobuild_am_traveltimes_formatted)
targets::tar_load(build_2050_am_traveltimes_formatted)
targets::tar_load(nobuild_pm_traveltimes_formatted)
targets::tar_load(build_2050_pm_traveltimes_formatted)

nobuild_am_traveltimes_formatted <- nobuild_am_traveltimes_formatted %>% mutate(Model = "No-Build", Label = paste0(Movement," (AM)"))
build_2050_am_traveltimes_formatted <- build_2050_am_traveltimes_formatted %>% mutate(Model = "Build", Label = paste0(Movement," (AM)"))
nobuild_pm_traveltimes_formatted <- nobuild_pm_traveltimes_formatted %>% mutate(Model = "No-Build", Label = paste0(Movement," (PM)"))
build_2050_pm_traveltimes_formatted <- build_2050_pm_traveltimes_formatted %>% mutate(Model = "Build", Label = paste0(Movement," (PM)"))

tt <- rbind(nobuild_am_traveltimes_formatted, build_2050_am_traveltimes_formatted, nobuild_pm_traveltimes_formatted, build_2050_pm_traveltimes_formatted) %>%
  select(Model, Movement, Label, TT = `Average Travel Time (sec)`, Distance = `Average Distance Traveled (ft)`) %>%
  arrange(Label)

tt$cont_y <- c(1, 1, 2, 2, 4, 4, 5, 5, 7, 7, 8, 8, 10, 10, 11, 11, 13, 13, 14, 14, 16, 16, 17, 17)

ggplot(tt) +
  geom_line(data = tt %>% filter(cont_y == 1), aes(x = TT, y = cont_y), color = "red") +
  geom_line(data = tt %>% filter(cont_y == 2), aes(x = TT, y = cont_y), color = "red") +
  geom_line(data = tt %>% filter(cont_y == 4), aes(x = TT, y = cont_y), color = "red") +
  geom_line(data = tt %>% filter(cont_y == 5), aes(x = TT, y = cont_y), color = "red") +
  geom_line(data = tt %>% filter(cont_y == 7), aes(x = TT, y = cont_y), color = "red") +
  geom_line(data = tt %>% filter(cont_y == 8), aes(x = TT, y = cont_y), color = "red") +
  geom_line(data = tt %>% filter(cont_y == 10), aes(x = TT, y = cont_y), color = "red") +
  geom_line(data = tt %>% filter(cont_y == 11), aes(x = TT, y = cont_y), color = "red") +
  geom_line(data = tt %>% filter(cont_y == 13), aes(x = TT, y = cont_y), color = "red") +
  geom_line(data = tt %>% filter(cont_y == 14), aes(x = TT, y = cont_y), color = "red") +
  geom_line(data = tt %>% filter(cont_y == 16), aes(x = TT, y = cont_y), color = "red") +
  geom_line(data = tt %>% filter(cont_y == 17), aes(x = TT, y = cont_y), color = "red") +
  geom_point(aes(x = TT, y = cont_y, color = Model), size = 2.5) +
  theme_bw() +
  labs(x = "Average Travel Time per Vehicle (seconds)",
       y = "Movement (Time of Day)") +
   scale_y_continuous(breaks = c(1, 2, 4, 5, 7, 8, 10, 11, 13, 14, 16, 17), 
                      labels = unique(tt$Label)) +
  scale_color_manual(values = c("Build" = "slateblue", "No-Build" = "red")) +
  scale_x_continuous(trans = log_trans(), breaks = c(seq(0,50,10), 75, seq(100,500,100), 15, 150), minor_breaks = c(seq(0,100,5), seq(100,500,50)), limits = c(10,NA), expand = expansion(c(0,0.05))) +
  theme(panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank())

```


## Conclusion

Under the No-Build intersection design, the model fails during the PM peak, exhibiting LOS F at all intersections.
These failing conditions are primarily due to the lack of designated turning lanes and severe overloading of the various intersections.
The build model performs to a high level of traffic management, with the highest LOS being exhibited at the I-15 SB interchange, performing at an LOS E. However, users should expect such delays due to the high traffic volume levels.
Providing for LOS C or better under such conditions would create a massive intersection, such as damage to safety or local ecology.

If such growth continues, additional highway interchanges would be necessary to provide additional capacity.
