# Conclusions and Recommendations

```{r}
#| label: setup
#| include: false

library(tidyverse)
library(kableExtra)
library(scales)
```

## Problem Statement

The City of Payson tasked the team to find an alternative design for the Payson Main Street I-15 interchange to handle future growth.
The team modeled existing traffic conditions using VISSIM, and by using MATSim modeling software, future growth rates over the next 27 years was predicted.
The team created future traffic models using VISSIM, and results were compiled and compared.
This section summarizes key results from the analysis.

## Key Results

Initial conditions provide limited traffic-based justifications for a redesign of the interchange.
However, due to a large number of intersection-related crashes, the team considered safety improvements as further justification for a redesign.
Additionally, if growth exceeds 30%, traffic conditions during the PM peak will be such that they will necessitate some traffic congestion mitigation efforts.

@tbl-final-intersection-comp compares the intersections from the Existing, 2050 No-Build, and 2050 Build conditions.

```{r}
#| label: tbl-final-intersection-comp
#| tbl-cap: Comparison of Intersection Delay and LOS

targets::tar_read(intersection_los_comp) %>% 
  mutate(
    time = str_to_upper(time),
    intersection = if_else(
      intersection == "SB Interchange",
      "SB Interchange*",
      intersection
    )
    ) %>% 
  kbl(align = "clccc", booktabs = TRUE, linesep = "",
      col.names = c(" ", "Intersection", "Existing", "No-Build", "Build")) %>% 
  # kable_styling(latex_options = "scale_down") %>%
  kable_styling() %>%
  column_spec(1, "1in") %>%
  column_spec(2, "1.5in") %>%
  # column_spec(3:5, "1in") %>%
  collapse_rows(columns = 1, valign = "middle") %>% 
  footnote(symbol = "This intersection is signalized in the Existing and No-Build models, but is unsignalized in the Build model.", threeparttable = TRUE)
```

The proposed redesign includes a dual-roundabout/dogbone interchange.
This design reduces speeds through the network and allows for safer pedestrian and bicycle access through the interchange.
Additionally, the design provides satisfactory traffic flow conditions compared to the heavy traffic volumes predicted in the No-Build scenario.
However, the existing conditions would also likely perform adequately by adding exclusive turning lanes and implementing a signal at the NB I-15 interchange.

## Recommendations

The team recommends the following future steps for the Payson Main Street. I-15 interchange:

-   Implementing a redesign to increase capacity at the interchange is necessary due to the predicted heavy growth.

-   The use of the provided design could provide a high-capacity solution that would also increase the safety of all users.

-   Implementing the design suggested here would require significant funds and construction time.
    The City of Payson should contract additional analysis on the addition of a signal at the NB I-15 interchange to coordinate with the SB I-15 interchange.
    This solution may be cheaper and have a shorter construction time.

-   Regardless, bridge expansion and reconstruction are likely necessary.
    The team highly recommends that the City of Payson and contracted design firms consider pedestrian and bicycle access during such rebuilding.
    Due to likely growth to the north of I-15, this may transform into a high pedestrian and bicycle access route between north and south Payson.

-   The additional WB leg to the NB interchange should either be removed and rerouted or added to a roundabout, as presented in this report, to remove unnecessary conflict points.
